---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  cache = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.height = "30%",
  fig.height = 8
)
```


# visPed R package

An R package that extends the kinship2 R package to plot pedigrees affected by multiple cancer syndromes. Designed to be compatible with the BayesMendel and associated packages.

## Installation

Use the latest github version by running

```{r, eval = FALSE}
devtools::install_github("bayesmendel/visPed")
library(visPed)
```

## Usage

Pass a compatible pedigree (data frame) into the function.

```{r, eval = FALSE}
visPed(pedigree)
```

To view the pedigree as a hierarchical network, which relies on the visNetwork package, run
```{r, eval = FALSE}
toVisNetwork(pedigree)
```

## Pedigree specifications

The user must specify the `pedigree` argument as a data frame with the following columns: 

- `ID`: A numeric value; ID for each individual. There should not be any duplicated entries. 
- `Sex`: A numeric value; `0` for female and `1` for male. Missing entries are not currently supported. 
- `MotherID`: A numeric value; unique ID for someone's mother. 
- `FatherID`: A numeric value; unique ID for someone's father. 
- `isProband`: A numeric value; `1` if someone is a proband, `0` otherwise. This will be overridden by the `proband` argument in `PanelPRO`, if it is specified. At least one proband should be specified by either the `isProband` column or `proband`. Multiple probands are supported. 
- `CurAge`: A numeric value; the age of censoring (current age if the person is alive or age of death if the person is dead). Ages ranging from `1` to `94` are allowed. 
- `isAffX`: A numeric value; the affection status of cancer `X`, where `X` is a `short` cancer code (see below). Affectation status should be encoded as `1` if the individual was diagnosed, `0 `otherwise. Missing entries are not currently supported. 
- `AgeX`: A numeric value; the age of diagnosis for cancer `X`, where `X` is a `short` cancer code (see below). Ages ranging from `1` to `94` are allowed. If the individual was not diagnosed for a given cancer, their affectation age should be encoded as `NA` and will be ignored otherwise. 
- `isDead`: A numeric value; `1` if someone is dead, `0` otherwise. Missing entries are assumed to be `0`. 
